Quantum Interval-Valued Probability: Measurement and Probability

### 2017-08-25 prop_iop.tex
prop_iop.tex \section{Expectation Values for IVPMs} Updates...

There are various kind of integrals defined for IVPMs, and the most
famous one is called the Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}.

\begin{definition}\label{def:Choquet}~
\begin{itemize}
\item Consider a classical IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\mul\left(E\right),\mur\left(E\right)\right]$.
\item We still decompose a random variable~$X$ into step functions, but
we separate the positive and negative terms and write like $\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}+\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}$,
i.e., 
\begin{equation}
x_{N^{-}}^{-}\le\ldots\le x_{1}^{-}\le0\le x_{1}^{+}\le\ldots\le x_{N^{+}}^{+}
\end{equation}
 and $\left\{ E_{i}^{-}\right\} _{i=1}^{N^{-}}$ and $\left\{ E_{i}^{+}\right\} _{i=1}^{N^{+}}$
are all disjoint.
\item Let $E_{i}^{s\prime}=\bigcup_{j\ge i}E_{j}^{s}$ and $\Delta\mu^{*}\left(E_{i}^{s\prime}\right)=\mu^{*}\left(E_{i}^{s\prime}\right)-\mu^{*}\left(E_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
\end{itemize}
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}
\int X\rmd\bar{\mu}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}x_{i}^{s}\left[\Delta\mul\left(E_{i}^{s\prime}\right),\Delta\mur\left(E_{i}^{s\prime}\right)\right]\,.\label{eq:interval-expectation}
\end{equation}
\end{definition}

@@

\begin{thm}\label{thm:Rosenmuller}\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}For
every classical IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ and
any random variable~$X:\Omega\rightarrow\mathbb{R}$, $\bar{\mu}$
is convex if and only if $\int X\rmd\bar{\mu}=\left[\min_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu,\max_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu\right]$.\end{thm}

@@

\yutsung{Put an example here if we need...}

@@

\begin{lemma}\label{lemma:deterministic-CIVPM}Given a CIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
the following statements are true:
\begin{enumerate}
\item Its left-end~$\mul$ and right-end~$\mur$ are the same CRVPM, and
will be simply denoted by $\pmeas:\events\rightarrow\left\{ 1,0\right\} $
\item There is a unique $\omega_{0}\in\Omega$ such that $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)=\left\{ \pmeas\left(\left\{ \omega_{0}\right\} \right)\right\} =\necess$,
denoted by $\pmeas^{-1}\left(1\right)$ or $\bar{\mu}^{-1}\left(\necess\right)$.
\end{enumerate}
\end{lemma}

>>>>

There is a subtlety happens when we want to define the expectation
value of a random variable on IVPMs, and will be demonstrated by the
following example.

\begin{example}\label{ex:expectation}Consider an CIVPM\nb$\bar{\mu}$
and an event\nb$E$ such that $\bar{\mu}(E)=[0.2,0.3]$ and $\bar{\mu}(\overline{E})=[0.7,0.8]$.
Given a random variable
\begin{equation} 
X\left(\omega\right)=\cases{ 
1 & , if $\omega\in E$ ;\\
2 & , if $\omega\notin E$ , 
}
\end{equation}
its expectation value with respect to $\bar{\mu}$ might be naïvely
defined as
\begin{equation}\eqalign{ 
 & 1\cdot\bar{\mu}(E)+2\cdot\bar{\mu}(\overline{E})\\
= & 1\cdot[0.2,0.3]+2\cdot[0.7,0.8]\\
= & [1\cdot0.2+2\cdot0.7,1\cdot0.3+2\cdot0.8]\\
= & [1.6,1.9]\,.
}\end{equation}
As we discussed when introducing the core, the true probability of
$\bar{\mu}(E)$, $\mu(E)$, can be anywhere in the range $[0.2,0.3]$
and the true probability of $\bar{\mu}(\overline{E})$ is $\mu(\overline{E})=1-\mu(E)$,
that is, $\mu\in\mathrm{core}\left(\bar{\mu}\right)$. If $\mu(E)=0.2$,
the expectation value of $X$ with respect to $\mu$ is 
\begin{equation}
1\cdot\mu(E)+2\cdot\mu(\overline{E})=1\cdot0.2+2\cdot0.8=1.8\,;
\end{equation}
if $\mu(E)=0.3$, the expectation value of $X$ with respect to $\mu$
is
\begin{equation}
1\cdot\mu(E)+2\cdot\mu(\overline{E})=1\cdot0.3+2\cdot0.7=1.7\,.
\end{equation}
In another word, no matter how we pick $\mu(E)$, the expectation
value of $X$ is in $\left[1.7,1.8\right]$ which is a more realistic
expectation value of $X$ with respect to $\bar{\mu}$. The interval~$\left[1.7,1.8\right]$
is narrower than the naïve expectation value~$[1.6,1.9]$ because
the naïve one combines contradicting situations in the computation.\end{example}

In general, the computation of the realistic expectation value, called
the Choquet integral, is more complex than what we discussed because
a random variable may have more than two values. Despite its complexity,
it is still a weight average among the values of a random variable
as given in the following definition.

\begin{definition}[Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}]\label{def:Choquet}~
\begin{itemize}
\item Consider a classical IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\mul\left(E\right),\mur\left(E\right)\right]$.
\item We still decompose a random variable~$X$ into step functions, but
we separate the positive and negative terms and write like $\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}+\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}$,
i.e., 
\begin{equation}
x_{N^{-}}^{-}\le\ldots\le x_{1}^{-}\le0\le x_{1}^{+}\le\ldots\le x_{N^{+}}^{+}
\end{equation}
 and $\left\{ E_{i}^{-}\right\} _{i=1}^{N^{-}}$ and $\left\{ E_{i}^{+}\right\} _{i=1}^{N^{+}}$
are all disjoint.
\item Let $E_{i}^{s\prime}=\bigcup_{j\ge i}E_{j}^{s}$ and $\Delta\mu^{*}\left(E_{i}^{s\prime}\right)=\mu^{*}\left(E_{i}^{s\prime}\right)-\mu^{*}\left(E_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
\end{itemize}
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}
\int X\rmd\bar{\mu}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}x_{i}^{s}\left[\Delta\mul\left(E_{i}^{s\prime}\right),\Delta\mur\left(E_{i}^{s\prime}\right)\right]\,.\label{eq:interval-expectation}
\end{equation}
\end{definition}

@@

\begin{thm}\label{thm:Rosenmuller}\cite{Rosenmuller1971,GilboaSchmeidler1994,Grabisch2016}For
every classical IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ and
any random variable~$X:\Omega\rightarrow\mathbb{R}$, $\bar{\mu}$
is convex if and only if
\begin{equation}
\int X\rmd\bar{\mu}=\left[\min_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu,\max_{\mu\in\mathrm{core}\left(\bar{\mu}\right)}\int X\rmd\mu\right]\,.\label{eq:Rosenmuller}
\end{equation}
\end{thm}

Notice that $\left[1.7,1.8\right]$ in example~\ref{ex:expectation}
is just the right-hand side of equation~(\ref{eq:Rosenmuller}).
Let compute the left-hand side of equation~(\ref{eq:Rosenmuller})
in table~\ref{table:Choquet} and see they are equal.
\begin{table}
\begin{tabular}{|c|c|c|c|}
\hline 
$i$ & 1 & 2 & 3\tabularnewline
\hline 
\hline 
$x_{i}^{+}$ & 1 & 2 & \tabularnewline
\hline 
$E_{i}^{+}$ & $E$ & $\overline{E}$ & \tabularnewline
\hline 
$E_{i}^{+\prime}$ & $E$ & $\Omega$ & $\emptyset$\tabularnewline
\hline 
$\mul\left(E_{i}^{+\prime}\right)$ &  &  & \tabularnewline
\hline 
 &  &  & \tabularnewline
\hline 
 &  &  & \tabularnewline
\hline 
 &  &  & \tabularnewline
\hline 
\end{tabular}\caption{\label{table:Choquet}}
\end{table}

@@

\subsection{Expectation Values for $0$-deterministic CIVPMs}

@@

\begin{lemma}\label{lemma:deterministic-CIVPM}Given a CIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
the following statements are true:
\begin{enumerate}
\item Its left-end~$\mul$ and right-end~$\mur$ are the same CRVPM, and
will be simply denoted by $\pmeas:\events\rightarrow\left\{ 1,0\right\} $.
\item There is a unique $\omega_{0}\in\Omega$ such that $\bar{\mu}\left(\left\{ \omega_{0}\right\} \right)=\left\{ \pmeas\left(\left\{ \omega_{0}\right\} \right)\right\} =\necess$,
denoted by $\pmeas^{-1}\left(1\right)$ or $\bar{\mu}^{-1}\left(\necess\right)$.
\end{enumerate}
\end{lemma}
